---
title: Relationship between prevalence of diarrhoea and underweight children in
  each county.
author: "Michelle Njeri Njuguna"
date: "2023-07-21"
output:
  word_document: default
  html_document:
    df_print: paged
---

# Import the required libraries

```{r}
library(tidyverse)
library(stats)
library(viridisLite)
library(lubridate)
library(visdat)
```

# Load the data to R from CSV

```{r}
input_data <- read.csv("cema_internship.csv", header = TRUE, sep = "," )
```

# Describe the data

```{r}
dim(input_data)    
glimpse(input_data)
summary(input_data)
```
# Missing Data
```{r}
# Convert columns with missing values to numeric using mutate_at()
cols_with_missing <- c("Acute_Malnutrition", "stunted_6_to_23_months", "stunted_0_to_less_than_6_months", "diarrhoea_cases",
                       "stunted_24_to_59_months","Total_Dewormed", "Underweight_0_to_less_6_months", 
                       "Underweight_6_to_23_months", "Underweight_24_to_59_Months")

input_data <- input_data %>%
  mutate_at(vars(all_of(cols_with_missing)), as.numeric)

# Visualize missing values 
vis_miss(input_data)
```
# Visualizing missing data so that we can see if the columns we want to use have sufficient data or they will bring about some inconsistency.

#Visualize your data (Total Dewormed in each county by Month)
```{r}
input_data$Month <- substr(input_data$period, 4, 6)

# Create Separate Bar Charts for Each County using split and map functions
unique_counties <- unique(input_data$county)

# Create a function to generate the bar chart
create_bar_chart <- function(county_data) {
  ggplot(county_data, aes(x = Month, y = Total_Dewormed, fill = period)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Total Dewormed in", unique(county_data$county), "by Month"),
         x = "Month",
         y = "Total Dewormed") +
    scale_y_continuous(labels = scales::comma) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(hjust = 0.5, margin = margin(t = 20, b = 40)))
}

# Split the data by county and create the bar charts using map function
county_data_list <- split(input_data, input_data$county)
bar_charts <- map(county_data_list, create_bar_chart)

# Print the bar charts
# This checks if de-worming curbs diarrhea
print(bar_charts)
```
# Diarrhoea Cases in each county by Month
```{r}
# Extract Month from "period" column
input_data$Month <- substr(input_data$period, 4, 6)

# Create Separate Bar Charts for Each County using split and map functions
unique_counties <- unique(input_data$county)

# Create a function to generate the bar chart
create_bar_chart <- function(county_data) {
  ggplot(county_data, aes(x = Month, y = diarrhoea_cases, fill = period)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Diarrhoea cases in", unique(county_data$county), "by Month"),
         x = "Month",
         y = "Diarrhoea cases") +
    scale_y_continuous(labels = scales::comma) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(hjust = 0.5, margin = margin(t = 20, b = 40)))
}

# Split the data by county and create the bar charts using map function
county_data_list <- split(input_data, input_data$county)
bar_charts <- map(county_data_list, create_bar_chart)

# Print the bar charts
print(bar_charts)
```
# Underweight_0_to_less_6_months
```{r}
# Extract Month from "period" column
input_data$Month <- substr(input_data$period, 4, 6)

# Create Separate Bar Charts for Each County using split and map functions
unique_counties <- unique(input_data$county)

# Create a function to generate the bar chart
create_bar_chart <- function(county_data) {
  ggplot(county_data, aes(x = Month, y = Underweight_0_to_less_6_months, fill = period)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Underweight 0 to < 6 months in", unique(county_data$county), "by Month"),
         x = "Month",
         y = "Underweight 0 to < 6 months") +
    scale_y_continuous(labels = scales::comma) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(hjust = 0.5, margin = margin(t = 20, b = 40)))
}

# Split the data by county and create the bar charts using map function
county_data_list <- split(input_data, input_data$county)
bar_charts <- map(county_data_list, create_bar_chart)

# Print the bar charts
print(bar_charts)
```
# Underweight 6 to 23 months
```{r}
# Extract Month from "period" column
input_data$Month <- substr(input_data$period, 4, 6)

# Create Separate Bar Charts for Each County using split and map functions
unique_counties <- unique(input_data$county)

# Create a function to generate the bar chart
create_bar_chart <- function(county_data) {
  ggplot(county_data, aes(x = Month, y = Underweight_6_to_23_months, fill = period)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Underweight 6 to 23 months in", unique(county_data$county), "by Month"),
         x = "Month",
         y = "Underweight 6 to 23 months") +
    scale_y_continuous(labels = scales::comma) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(hjust = 0.5, margin = margin(t = 20, b = 40)))
}

# Split the data by county and create the bar charts using map function
county_data_list <- split(input_data, input_data$county)
bar_charts <- map(county_data_list, create_bar_chart)

# Print the bar charts
print(bar_charts)
```

 
# Underweight 24 to 59 Months
```{r}
# Extract Month from "period" column
input_data$Month <- substr(input_data$period, 4, 6)

# Create Separate Bar Charts for Each County using split and map functions
unique_counties <- unique(input_data$county)

# Create a function to generate the bar chart
create_bar_chart <- function(county_data) {
  ggplot(county_data, aes(x = Month, y = Underweight_24_to_59_Months, fill = period)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Underweight 24 to 59 Months in", unique(county_data$county), "by Month"),
         x = "Month",
         y = "Underweight 24 to 59 Months") +
    scale_y_continuous(labels = scales::comma) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          plot.title = element_text(hjust = 0.5, margin = margin(t = 20, b = 40)))
}

# Split the data by county and create the bar charts using map function
county_data_list <- split(input_data, input_data$county)
bar_charts <- map(county_data_list, create_bar_chart)

# Print the bar charts
print(bar_charts)
```
# Correlate Total Dewormed with Diarrhoea Cases in each county
```{r}
# Calculate Correlation between Total Dewormed and Diarrhoea Cases for Each County
correlation_by_county <- input_data %>%
  group_by(county) %>%
  summarize(correlation_coefficient = cor(Total_Dewormed, `diarrhoea_cases`, use = "pairwise.complete.obs"),
            correlation_percentage = correlation_coefficient * 100) %>% arrange(desc(correlation_coefficient)) 

print(correlation_by_county)
```
# Correlate Diarrhoea Cases with Underweight 0-<6 months in each county
```{r}
# Calculate Correlation between Diarrhoea Cases and Underweight 0-<6 months for Each County
correlation_by_county <- input_data %>%
  group_by(county) %>%
  summarize(correlation_coefficient = cor(diarrhoea_cases, `Underweight_0_to_less_6_months`, use = "pairwise.complete.obs"),
            correlation_percentage = correlation_coefficient * 100) %>% arrange(desc(correlation_coefficient))

print(correlation_by_county)
```
# Correlate Diarrhoea Cases with Underweight 6 to 23 months in each county
```{r}
# Calculate Correlation between Diarrhoea Cases and Underweight 6 to 23 months for Each County
correlation_by_county <- input_data %>%
  group_by(county) %>%
  summarize(correlation_coefficient = cor(diarrhoea_cases, `Underweight_6_to_23_months`, use = "pairwise.complete.obs"),
            correlation_percentage = correlation_coefficient * 100) %>% arrange(desc(correlation_coefficient))

print(correlation_by_county)
```
# Correlate Diarrhoea Cases with Underweight 24 to 59 months in each county
```{r}
# Calculate Correlation between Diarrhoea Cases and Underweight 24 to 59 months for Each County
correlation_by_county <- input_data %>%
  group_by(county) %>%
  summarize(correlation_coefficient = cor(diarrhoea_cases, `Underweight_24_to_59_Months`, use = "pairwise.complete.obs"),
            correlation_percentage = correlation_coefficient * 100) %>% arrange(desc(correlation_coefficient))

print(correlation_by_county)
```
# NOTE!
# If the correlation is greater than 70%, it shows that if a child suffers from diarrhoea there's a great possibility the child may also be underweight.
# If the correlation is ranging from 40-60%, it shows that if a child suffers from diarrhoea there may be a possibility of them being underweight.
# If the correlation is less than 40%, it shows that if a child suffers from diarrhoea there's a small possibility of them being underweight.
# Correlation percentages are ordered from the highest to the lowest county.


